---
title: "portfoliotools-guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{portfolioTools-guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

# load libraries
library(portfoliotools)
library(dplyr)

```

# portfolioTools
portfoliotools is a small R package that provides simple helper functions
for data cleaning, filtering, summarising, and selecting columns.
It is designed for beginner-friendly and consistent data manipulation workflows.

# Installation
Install the package from GitHub:


``` r

devtools::install_github("RachelvanderKreeft/portfoliotools")

```

# Example dataset

We will use a simple dataset:

```{r}

df <- data.frame(
  age = c(20, 30, 40, NA),
  score = c(10, 20, NA, 40),
  group = c("A", "A", "B", "B")
)

df

```


## 1. Filter values in a range
filter_range() keeps rows where a numeric column falls within a range.

```{r}

filter_range(df, age, 20, 40)

```


## 2. Remove rows with missing values
drop_na_cols() removes rows where selected columns contain NA values.

```{r}

drop_na_cols(df, age, score)

```


## 3. Select specific columns
select_columns() keeps only selected columns.

```{r}

select_columns(df, age, score)

```


## 4. Summarise by group
summary_by_group() calculates mean, standard deviation and number of observations per group.

```{r}

summary_by_group(df, group, score)

```


## Workflow example
These functions can be combined into a simple data pipeline:

```{r}

df %>%
  filter_range(age, 20, 40) %>%
  drop_na_cols(age, score) %>%
  select_columns(age, score)

```


## Summary

portfolioTools provides a simple toolkit for: \
-filtering data \
-cleaning missing values \
-selecting columns \
-summarising grouped data \
\
It is designed to support readable and consistent data analysis workflows in R. \
